---
title: "ST558HW2_BaseRprogramming"
format: html
editor: visual
---

## Task 1: Vectors

Create vectors for pre- and post-treatment blood pressure measurements:

```{r}
pre_bp <- c(120, 151, 125, 126, 115, 132, 132, 129, 134, 139, 127, 122, 127, 135, 133, 128, 147, 138, 140, 132)
post_bp <- c(127, 145, 135, 122, 115, 122, 123, 126, 126, 129, 132, 146, 120, 114, 121, 120, 128, 120, 133, 115)
```

Generate generic patient names

```{r}
patient <-paste("Subject", 1:20, sep = "_")
print(patient)

```

Assign names to both vectors

```{r}
names(pre_bp) <- patient
names(post_bp)<- patient
print(pre_bp)
print(post_bp)
```

Calculate change in blood pressure by subtracting post-treatment from pre-treatment measurement, then calculate the average DECREASE (or change) in blood pressure *across all patients*.

```{r}
diff_bp <- pre_bp - post_bp
print(mean(diff_bp))
```

Identify patients (by generic name and index) who experienced a decrease in blood pressure after treatment.

```{r}

#decrBP <- which(BPchange > 0)
#print(decrBP)
print(which(diff_bp > 0))


```

Subset the vector of differences to only return those that have a positive change.

```{r}


decr_bp <- diff_bp[diff_bp > 0]
print(decr_bp)
```

Calculate the average decrease in blood pressure for those whose blood pressure decreased.

```{r}
print(mean(decr_bp))
```

## Task 2: Data Frames

Create a data frame object with four columns corresponding to your data above: patient, pre_bp, post_bp, and diff_bp.

```{r}
treatment_df <- data.frame(patient, pre_bp, post_bp, diff_bp, row.names = NULL)

treatment_df
```

Return only rows of the data frame where the diff_bp column is negative.

```{r}

#df[df$x > 1, ]
treatment_df[treatment_df$diff_bp < 0,]

```

Add a new column to the data frame corresponding to TRUE if the post_bp is less than 120 and FALSE if it is not.

```{r}

treatment_df$PostLess120 <- treatment_df$post_bp < 120
treatment_df
```

Finally, print the data frame out nicely.

```{r}

knitr::kable(treatment_df)

```

## Task 3: List Practice

```{r}
placebo_pre_bp <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
placebo_post_bp <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)
placebo_diff <- placebo_pre_bp - placebo_post_bp
patient <-paste("Subject", 1:10, sep = "_")

placebo_df <- data.frame(patient, placebo_pre_bp, placebo_post_bp, placebo_diff)

placebo_df$placebo_PostLess120 <-placebo_df$placebo_post_bp < 120

placebo_df
```

Create and store a list with two elements: element 1 = treatment, and element 2 = placebo, then name the elements of the list.

```{r}

my_list <- list(treatment_df, placebo_df)
my_list <- list(treatment = treatment_df, placebo = placebo_df)

```

Access the first list element using three different types of syntax (the result can be a list or just the data frame).

```{r}
my_list[1]
str(my_list[1])
my_list$treatment

```

Use your list from question 2. In one line of code, access the pre_bp column of the placebo data frame.

```{r}

my_list[[2]][2]

```
